---
title: "MATH60633A – TP1"
output: html_document
date: "2025-02-15"
---
Estimate and backtest the value-at-risk (VaR) of a portfolio of two indices using the
GARCH(1,1)-Normal model.


### Load libraries and process data 


```{r, message=FALSE, warning=FALSE}
library(PerformanceAnalytics)
library(GAS)
library(here)

source(here('src','core_functions.R'))
load('data/indices.rda')

# Consider only index values since January 2005
prices <- prices['2005-01-01/']

# Compute the log-returns for both series
logret <-  CalculateReturns(prices, 'log')[-1]

```

### Plot cumsum logrets 


```{r, message=FALSE, warning=FALSE}

par(mfrow = c(2, 1))
plot(cumsum(logret[, 1]), 
     col = 'orange', 
     xlab = 'Time', 
     ylab = 'Cumsum Log-returns',
     main = 'S&P500')
plot(cumsum(logret[, 2]),
     col = 'blue', 
     xlab = 'Time', 
     ylab = 'Cumsum Log-returns',
     main = 'FTSE100')

```

## Static estimation of the 95% VaR

```{r, message=FALSE, warning=FALSE}


T = 1000
level = 0.95

## GARCH
forecast_SP500 <- f_forecast_var(as.numeric(logret[1:T, 1]), level)
forecast_FTSE <- f_forecast_var(as.numeric(logret[1:T, 2]), level)

# VaR forecast at T+1 
VaR_SP500 <- forecast_SP500$VaR_Forecast
VaR_FTSE100 <- forecast_FTSE$VaR_Forecast

# MLE GARCH parameters
GARCH_param_SP500 <- forecast_SP500$GARCH_param
GARCH_param_FTSE <- forecast_FTSE$GARCH_param
```

Parameter estimation results

```{r, message=FALSE, warning=FALSE}

# Create a data frame with Omega, Alpha, and Beta 
GARCH_table <- data.frame(
  Index = c("SP500", "FTSE"),
  Omega = format(round(c(GARCH_param_SP500[1], GARCH_param_FTSE[1]), 10), nsmall = 10, scientific = TRUE),
  Alpha = format(round(c(GARCH_param_SP500[2], GARCH_param_FTSE[2]), 3), nsmall = 3, scientific = FALSE),
  Beta  = format(round(c(GARCH_param_SP500[3], GARCH_param_FTSE[3]), 3), nsmall = 3, scientific = FALSE)
)

GARCH_table



```
The VaR forecast for T+1 indicates that the S&P 500 carries greater risk than the FTSE 100 based on this metric.

```{r, message=FALSE, warning=FALSE}
VaR_forecast <- data.frame(
  Index = c("VaR_forecast"),
  SP500 = VaR_SP500,
  FTSE = VaR_FTSE100
)

VaR_forecast
```

## Backtesting


```{r, message=FALSE, warning=FALSE}

roll_garch_SP500 <- f_roll_forecast_var(y = logret[1:(2*T), 1], 
                         level = level, window_size = T, n_forecast = T)

roll_garch_FTSE100 <- f_roll_forecast_var(y = logret[1:(2*T), 2], 
                         level = level, window_size = T, n_forecast = T)

```


### Plots


```{r, message=FALSE, warning=FALSE}
# Create df to plot
df <- xts(
  data.frame(logret[(T + 1):(2*T), ],
              var_SP500 = roll_garch_SP500$roll_VaR_Forecast, 
              var_FTSE100 = roll_garch_FTSE100$roll_VaR_Forecast,
              cvar_SP500 = roll_garch_SP500$roll_ConditionalVariances,
              cvar_FTSE100 = roll_garch_FTSE100$roll_ConditionalVariances),
  order.by = index(logret)[(T + 1):(2*T)])

par(mfrow = c(1, 1))
matplot(df[, 3:4], 
     col = c("orange", "blue"), 
     type = "l",  
     lty = 1, 
     lwd = 2,
     xlab = 'Time', 
     ylab = 'VaR', 
     main = 'VaR S&P500 & FTSE100')
grid(col = "gray", lty = "dotted", lwd = 1)
legend("bottomright", 
       legend = c("S&P 500", "FTSE 100"), 
       col = c("orange", "blue"), 
       lty = 1, 
       cex = 0.8,
       lwd = 2) 


```


### VaR and the returns for the S&P500 in the same plot


```{r, message=FALSE, warning=FALSE}
plot(index(df),
     as.numeric(df$SP500), 
     type = 'l', 
     col = 'blue', 
     xlab = 'Time', 
     ylab = 'Returns', 
     main = 'S&P500 - Rolling Window VaR & Returns')
lines(index(df), 
      as.numeric(df$var_SP500), 
      col = 'red',
      lwd = 2)
grid(col = "gray", lty = "dotted", lwd = 1)
legend("topright", 
       legend = c("Returns", "Predicted VaR"),
       col = c("blue", "red"), 
       lty = 1:1, cex = 0.8, lwd = 2)

```


### VaR and the returns for the FTSE100 in the same plot

```{r, message=FALSE, warning=FALSE}

plot(index(df), 
     as.numeric(df$FTSE100), 
     type = 'l', 
     col = 'blue', 
     xlab = 'Time', 
     ylab = 'Returns', 
     main = 'FTSE100 - Rolling Window VaR & Returns')
lines(index(df), 
      as.numeric(df$var_FTSE100), 
      col = 'red',
      lwd = 2)
grid(col = "gray", lty = "dotted", lwd = 1)
legend("topright", 
       legend = c("Returns", "Predicted sVaR"),
       col = c("blue", "red"), 
       lty = 1:1, cex = 0.8, lwd = 2)

```
#### Coverage raio

```{r, message=FALSE, warning=FALSE}

violations_SP500 <- sum(df$SP500 < df$var_SP500)
violations_FTSE100 <- sum(df$FTSE100 < df$var_FTSE100)

data.frame(
  Index = "Coverage",
  SP500 =  violations_SP500/((1-level)*length(df$var_SP500)),
  FTSE100 =  violations_FTSE100/((1-level)*length(df$var_FTSE100))
)

```



## Bonus: Kupiec and  Christoffersen tests

```{r, message=FALSE, warning=FALSE}

# GAS package: backtest and formal tests

GAS_backtest_SP500 <- GAS::BacktestVaR(data = as.numeric(df$SP500),
                                        VaR = as.numeric(df$var_SP500),
                                         alpha = .05)

GAS_backtest_FTSE100 <- GAS::BacktestVaR(data = as.numeric(df$FTSE100),
                                       VaR = as.numeric(df$var_FTSE100),
                                        alpha = .05)

GAS_test_results <- data.frame(
  Index = rep(c("SP500", "FTSE100"), each = 2),
  Test = rep(c("LRuc", "LRcc"), times = 2),
  Test_Value = c(GAS_backtest_SP500$LRuc["Test"], GAS_backtest_SP500$LRcc["Test"], 
                 GAS_backtest_FTSE100$LRuc["Test"], GAS_backtest_FTSE100$LRcc["Test"]),
  P_Value = c(GAS_backtest_SP500$LRuc["Pvalue"], GAS_backtest_SP500$LRcc["Pvalue"], 
              GAS_backtest_FTSE100$LRuc["Pvalue"], GAS_backtest_FTSE100$LRcc["Pvalue"])
)

GAS_test_results
```


















